---
title: "Panel Data Analysis"
author: "Zahid Asghar"
date: "4/8/2020"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(stargazer)
library(tidyverse)


```

## Panel Data


```{r Fatalities, warning=FALSE,message=FALSE}
library(AER)
library(dplyr)
data("Fatalities")
#glimpse(Fatalities)
Fatalities %>% group_by(year) %>% summarise(number=n())
```

## Manipulaton
$$ FatalityRate_{it} = \beta_0 + \beta_1 BeerTax_{it} + \beta_2 Z_{i} + u_{it}$$
$$\widehat{FatalityRate_{i1988} - FatalityRate_{i1982}} = -\underset{(0.065)}{0.072} -\underset{(0.36)}{1.04} \times (BeerTax_{i1988}-BeerTax_{i1982}).$$
?plm
You can also embed plots, for example:
```{r}
Fatalities<-Fatalities %>% mutate(fatal_rate=fatal/pop*10000)
fat1982<-Fatalities %>% filter(year==1982)
fat1988<-Fatalities %>% filter(year==1988)

fat1982_mod<-lm(fatal_rate~beertax,data = fat1982)
fat1988_mod<- lm(fatal_rate~beertax,data = fat1988)
library(moderndive)
library(stargazer)
get_regression_table(fat1982_mod)
get_regression_table(fat1988_mod)
stargazer(fat1982_mod,fat1988_mod,type="text")

```


```{r}
library(ggplot2)
ggplot(fat1982)+aes(x=beertax,y=fatal_rate)+geom_point()+geom_smooth(method = "lm",se=FALSE)+
  labs(x = "Beer tax (in 1982 dollars)",
     y= "Fatality rate (fatalities per 10000)",
     title = "Traffic Fatality Rates and Beer Taxes in 1982")+
     ylim(0,4.5)

ggplot(fat1988)+aes(x=beertax,y=fatal_rate)+geom_point()+geom_smooth(method = "lm",se=FALSE)+
  labs(x = "Beer tax (in 1988 dollars)",
     y= "Fatality rate (fatalities per 10000)",
     title = "Traffic Fatality Rates and Beer Taxes in 1982")+
     ylim(0,4.5)


```



## Panel with two time periods


```{r}
# compute the differences 
diff_fatal_rate <- fat1988$fatal_rate - fat1982$fatal_rate
diff_beertax <- fat1988$beertax - fat1982$beertax
df<-data.frame(diff_fatal_rate,diff_beertax)
# estimate a regression using differenced data
fatal_diff_mod <- lm(diff_fatal_rate ~ diff_beertax)

ggplot(df)+aes(x=diff_beertax,y=diff_fatal_rate)+geom_point()+geom_smooth(method = "lm",se=FALSE)+
  labs(x = "Change in beer tax (in 1988 dollars)",
     y = "Change in fatality rate (fatalities per 10000)",
     title = "Changes in Traffic Fatality Rates and Beer Taxes in 1982-1988")+
     xlim (-0.6,0.6)+ ylim (-1.5,1)

```


## Fixed Effect Regression

$$Y_{it} = \beta_0 + \beta_1 X_{it} + \beta_2 Z_i +  u_{it}$$
$$\begin{align}
FatalityRate_{it} = \beta_1 BeerTax_{it} + StateFixedEffects + u_{it}, \tag{10.6}
\end{align}$$


```{r}
library(plm)
# obtain demeaned data
Fatalities_demeaned <- Fatalities %>%group_by(state) %>% 
  mutate(fatal_rate=fatal_rate-mean(fatal_rate),beertax=beertax-mean(beertax))
  
  
  
  
# estimate the regression
summary(lm(fatal_rate ~ beertax - 1, data = Fatalities_demeaned))

# estimate the fixed effects regression with plm()
fatal_fe_mod <- plm(fatal_rate ~ beertax, 
                    data = Fatalities,
                    index = c("state", "year"), 
                    model = "within")

# print summary using robust standard errors
coeftest(fatal_fe_mod, vcov. = vcovHC, type = "HC1")





```


$$\begin{align}
\widehat{FatalityRate} = -\underset{(0.29)}{0.66} \times BeerTax + StateFixedEffects. \tag{10.7}
\end{align}$$

### Regression with Time Fixed Effects
$$Y_{it} = \beta_0 + \beta_1 X_{it} + \delta_2 B2_t + \cdots + \delta_T BT_t + u_{it}$$
```{r}
#Via regression method
fatal_tefe_lm_mod <- lm(fatal_rate ~ beertax + state + year - 1, data = Fatalities)
fatal_tefe_lm_mod
# via plm()
fatal_tefe_mod <- plm(fatal_rate ~ beertax, 
                      data = Fatalities,
                      index = c("state", "year"), 
                      model = "within", 
                      effect = "twoways")

coeftest(fatal_tefe_mod, vcov = vcovHC, type = "HC1")

```


State and Year variables are factors and lm() function generates dummies for these factor variables n + (T-1) = 48 + 6 = 54
while plm() function just provide estimate of beertax.
$$\begin{align}
\widehat{FatalityRate} =  -\underset{(0.35)}{0.64} \times BeerTax + StateEffects + TimeFixedEffects. \tag{10.8}
\end{align}$$

$$Y_{it} = \beta_1 X_{it} + \alpha_i + u_{it} \ \ , \ \ i=1,\dots,n, \ t=1,\dots,T,$$

```{r}
# obtain a summary based on heteroskedasticity-robust standard errors 
# (no adjustment for heteroskedasticity only)
coeftest(fatal_tefe_lm_mod, vcov = vcovHC, type = "HC1")[1, ]
# check class of the (plm) model object
class(fatal_tefe_mod)
## [1] "plm"        "panelmodel"
```


##




```{r}
library(stargazer)
# discretize the minimum legal drinking age
Fatalities$drinkagec <- cut(Fatalities$drinkage,
                            breaks = 18:22, 
                            include.lowest = TRUE, 
                            right = FALSE)

# set minimum drinking age [21, 22] to be the baseline level
Fatalities$drinkagec <- relevel(Fatalities$drinkagec, "[21,22]")

# mandadory jail or community service?
Fatalities$punish <- with(Fatalities, factor(jail == "yes" | service == "yes", 
                                             labels = c("no", "yes")))

# the set of observations on all variables for 1982 and 1988
Fatalities_1982_1988 <- Fatalities[with(Fatalities, year == 1982 | year == 1988), ]
```
Next, we estimate all seven models using plm().
```{r}
# estimate all seven models
fatalities_mod1 <- lm(fatal_rate ~ beertax, data = Fatalities)

fatalities_mod2 <- plm(fatal_rate ~ beertax + state, data = Fatalities)

fatalities_mod3 <- plm(fatal_rate ~ beertax + state + year,
                       index = c("state","year"),
                       model = "within",
                       effect = "twoways", 
                       data = Fatalities)

fatalities_mod4 <- plm(fatal_rate ~ beertax + state + year + drinkagec 
                       + punish + miles + unemp + log(income), 
                       index = c("state", "year"),
                       model = "within",
                       effect = "twoways",
                       data = Fatalities)

fatalities_mod5 <- plm(fatal_rate ~ beertax + state + year + drinkagec 
                       + punish + miles,
                       index = c("state", "year"),
                       model = "within",
                       effect = "twoways",
                       data = Fatalities)

fatalities_mod6 <- plm(fatal_rate ~ beertax + year + drinkage 
                       + punish + miles + unemp + log(income), 
                       index = c("state", "year"),
                       model = "within",
                       effect = "twoways",
                       data = Fatalities)

fatalities_mod7 <- plm(fatal_rate ~ beertax + state + year + drinkagec 
                       + punish + miles + unemp + log(income), 
                       index = c("state", "year"),
                       model = "within",
                       effect = "twoways",
                       data = Fatalities_1982_1988)
```

We again use stargazer() (Hlavac, 2018) to generate a comprehensive tabular presentation of the results.
---
output: pdf_document
---

```{r }
library(stargazer)
# gather clustered standard errors in a list
rob_se <- list(sqrt(diag(vcovHC(fatalities_mod1, type = "HC1"))),
               sqrt(diag(vcovHC(fatalities_mod2, type = "HC1"))),
               sqrt(diag(vcovHC(fatalities_mod3, type = "HC1"))),
               sqrt(diag(vcovHC(fatalities_mod4, type = "HC1"))),
               sqrt(diag(vcovHC(fatalities_mod5, type = "HC1"))),
               sqrt(diag(vcovHC(fatalities_mod6, type = "HC1"))),
               sqrt(diag(vcovHC(fatalities_mod7, type = "HC1"))))
library(gt)

library(huxtable)
huxreg(fatalities_mod1, fatalities_mod2, fatalities_mod3, 
          fatalities_mod4, fatalities_mod5, fatalities_mod6, fatalities_mod7)
# generate the table
```

